

\fixme{I'm not sure about the I versus We language here.  What's
  appropriate for an abstract?  I looked up an example online and that
  person was using 'We'.} This thesis first reports on computational
research in two different areas.  I first discuss the Min-protein
system found within \emph{Escherichia coli}.  Followign this I discuss
an investigation into improving free energy functionals that are used
within Classical Density Functional Theory in order to model water
%
We begin by examining the dynamics of the Min-protein system within
E. Coli, which aid in regulating the process of cell division by
identifying the center of the cell.  While this system usually
exhibits robust bipolar oscillations in a variety of cell shapes,
recent experiments have shown that when the cells are mechanically
deformed into wide, flattened out, irregular shapes, the spatial
regularity of these oscillations breaks down. We employ widely used
stochastic and deterministic models of the Min system to simulate
cells with flattened shapes.  The deterministic model predicts strong
bipolar oscillations, in contradiction with the experimentally
observed behavior, while the stochastic model, which is based on the
same reaction-diffusion equations, does predict spatially irregular
oscillations as observed in experiment.  We further report simulations
of symmetric but flattened cell shapes, and find that it is the
flattening rather than the asymmetry of the cell shapes that causes
the irregular oscillation behavior.
%

We begin our discussion of our Classical Density Functional Theory
research by investigating the value of the distribution function of an
inhomogeneous hard-sphere fluid at contact. This quantity plays a
critical role in statistical associating fluid theory, which is the
basis of a number of recently developed classical density functionals,
including ones developed within my research group. We define two
averaged values for the distribution function at contact and derive
formulas for each of them from the White Bear version of the
fundamental measure theory functional, using an assumption of
thermodynamic consistency. We test these formulas, as well as two
existing formulas, against Monte Carlo simulations and find excellent
agreement between the Monte Carlo data and one of our averaged
distribution functions.
%

We then modify our recently published statistical associating fluid
theory-based classical density functional theory for water,
incorporating this improved distribution function at contact.  We
examine the effect of this alteration by studying two hard-sphere
solutes and a Leonard jones approximation of a krypton-atom solute,
and find improvement.
%% We then modify our recently published statistical associating fluid
%% theory-based classical density functional theory for water,
%% incorporating this improved distribution function at contact.  Our
%% function replaces another in the association term of our free energy,
%% improving its description of hydrogen bonding. We examine the effect
%% of this improvement by studying two hard solutes (a hard hydrophobic
%% rod and a hard sphere) and a Lennard-Jones approximation of a krypton
%% atom solute. The improved functional leads to a moderate change in the
%% density profile and a large decrease in the number of hydrogen bonds
%% broken in the vicinity of the hard solutes. We find an improvement of
%% the partial radial distribution for a krypton atom in water when
%% compared with experiment.
%

Finally, we introduce an approximation for the pair distribution
function of the inhomogeneous hard sphere fluid. Our approximation
makes use of the new distribution function at contact refered to
above. This approach achieves greater computational efficiency than
previous approaches by enabling the use of exclusively fixed-kernel
convolutions, which allows for an implementation using fast Fourier
transforms. We compare results for our pair distribution approximation
with two previously published works and Monte Carlo simulation,
showing favorable results.



